A fully labelled proof system for intuitionistic modal logics

### 2019-10-08 caleidoscope/conclusion.tex
final pass

In this paper we embrace the fully labelled approach to intuitionistic modal logic as pioneered by~\cite{maffezioli:naibo:negri} and generalise it to the class of logics defined by (one-sided intuitionistic) Scott-Lemmon axioms.
%
We establish that it is a valid approach to intuitionistic modal logic by proving soundness and completeness of our system, via a reductive cut-elimination argument.

We believe that it is furthermore the more appropriate way to treat logics outside of the path axioms definable fragment.
%
However, we have not showed that our system satisfies Simpson's 6th requirement, that is, "there is an intuitionistically comprehensible explanation of the meaning of the modalities relative to which [our system] is sound and complete".
%
To make sure that his system satisfies this requirement, Simpson chose to depart from the direct correspondence with modal axioms and their corresponding class of Kripke frames.
%
We instead took the way of a direct correspondence of our system with the class of frames defined by Scott-Lemmon axioms as uncovered by~\cite{plotkin:sterling}, but as this class of logics seems to be rather well-behaved, we believe it shoud be possible to prove the satisfaction of Simpson's 6th requirement too.

>>>>

In this paper we embrace the fully labelled approach to intuitionistic modal logic as pioneered by~\cite{maffezioli:etal:synthese13} and generalise it to the class of logics defined by (one-sided intuitionistic) Scott-Lemmon axioms.
%
We establish that it is a valid approach to intuitionistic modal logic by proving soundness and completeness of our system, via a reductive cut-elimination argument.

For a restricted class of logics defined by so-called \emph{path axioms}: $(\DIA^k\BOX A \IMP\BOX^m A) \AND (\DIA^k A \IMP \BOX^m\DIA A)$ the standard labelled framework with one relation $R$ was enough for Simpson to get a strong connection between the sequent system, the axiomatisation and the birelational semantics~\ref{simpson:phd}.
%
%\todo{For path axioms we must have that adding the rule for the two sides of the axiom is equivalent to adding the rule that doesn't make use of the pre-order. Is it possible to see that directly on the proof theoretical side?}
%\sonia{probably not because the preorder is introduced/used in other rules for $\BOX$ and $\IMP$}
%
We believe that the framework presented here might be the more appropriate way to treat logics outside of the path axioms definable fragment.

However, we have not showed that our system satisfies Simpson's 6th requirement, that is, "there is an intuitionistically comprehensible explanation of the meaning of the modalities relative to which [our system] is sound and complete".
%
To make sure that his system satisfies this requirement, Simpson chose to depart from the direct correspondence with modal axioms and their corresponding class of Kripke frames, and to study intuitionistic purely as a fragment of intuitionistic first-order logic.
%
We instead took the way of a direct correspondence of our system with the class of frames defined by Scott-Lemmon axioms as uncovered by~\cite{plotkin:stirling:86}, but as this class of logics seems to be rather well-behaved, we believe it shoud be possible to prove the satisfaction of Simpson's 6th requirement too.

### 2019-10-08 caleidoscope/system.tex
final pass

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.
%
These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

@@

	\item $\fm{\DIA A}$
%	\begin{smallequation*}
%		\vlinf{\labrn{id_g}}{}{\B, x \le y; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{}
%		\reducesto
%	\end{smallequation*}
	\begin{smallequation*}
		\vlderivation{
			\vlin{\llabrn\DIA}{}{\B, \futs xy, \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{
				\vlin{\color{red}\rn{F_2}}{}{\B, \futs xy, \accs xz, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
					\vlin{\rlabrn\DIA}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
						\vlin{\labrn{id_g}}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}, \labels{u}{A}}{
							\vlhy{}
						}
					}
				}
			}
		}
	\end{smallequation*}
        \qedhere
	\end{itemize}
\end{proof}

>>>>

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
%\sonia{This is an interesting alternative view but for me the starting point really is the Maffezioli, Naibo and Negri paper. Maybe rephrase slightly to acknowledge their work too.}
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{str:2017maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this naive translation would also yield the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A \IMP B} \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, \accs xz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, \accs xz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{str:2017maehara}, these are the two rules that force
single-conclusion. 
%This is the reason why in our system we need the
%$\le$ in the relational atoms; we need to go to the future of the
%current world.  
%
In our system, this phenomenon is replaced by a re-positioning of the considered formulas to a fresh label.
%
In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that make use of the
pre-order relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.
%
These can be obtained by applying the axioms-as-rules methodology as in~\cite{maffezioli:etal:synthese13}.

@@

	\item $\fm{\DIA A}$
%	\begin{smallequation*}
%		\vlinf{\labrn{id_g}}{}{\B, x \le y; \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{}
%		\reducesto
%	\end{smallequation*}
	\begin{smallequation*}
		\vlderivation{
			\vlin{\llabrn\DIA}{\lb z \mbox{ fresh}}{\B, \futs xy, \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{
				\vlin{\color{red}\rn{F_2}}{\lb u \mbox{ fresh}}{\B, \futs xy, \accs xz, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
					\vlin{\rlabrn\DIA}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
						\vlin{\labrn{id_g}}{}{\B, \futs xy, \accs xz, \futs zu, \accs yu, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}, \labels{u}{A}}{
							\vlhy{}
						}
					}
				}
			}
		}
	\end{smallequation*}
        \qedhere
	\end{itemize}
\end{proof}
